---
title: "Example analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, error=FALSE,message=FALSE}
library(matrixStats)
library(tidyverse)
library(ggridges)
```

# 1B Customize the website

# 1C: Example data:

## what is the data

The data that I chose for this project is about childcare costs from 2008-2018 from [National Database of Childcare Prices](https://www.dol.gov/agencies/wb/topics/featured-childcare). My goal is to understand how the childcare cost in Maryland has changed during the years and how it varies across counties. I will perform this analysis for different childcare provider type and children ages. Data dictionary can be found [here](https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-05-09/readme.md).

### Load in the data in R

```{r, warning=FALSE, error=FALSE}
library(tidytuesdayR)
#Check if the directory exist:
if(!dir.exists("../data")){
 dir.create("../data")
 #only download data once:
 if(!file.exists(c("../data/childcare_costs.rda","../data/counties.rda"))){
 tuesdata <- tidytuesdayR::tt_load('2023-05-09')
tuesdata <- tidytuesdayR::tt_load(2023, week = 19)

childcare_costs <- tuesdata$childcare_costs
counties <- tuesdata$counties
 # save the files like this:
 library(usethis)
 use_data(childcare_costs)
 use_data(counties)
 }}
```

```{r, warning=FALSE, error=FALSE}
load("../data/childcare_costs.rda")
load("../data/counties.rda")
```

## Data wrangling
First join the childcare costs with county data
Only select data for Maryland after removing the NA values

```{r, warning=FALSE, error=FALSE}
childcare<-inner_join(childcare_costs,counties) %>%
  drop_na() %>% 
  filter(state_abbreviation=="MD")
```


#1 which childcare provider is more expensive
To answer this question, I will plot (geom_smooth) the cost of childcare during years colored by different facilities.

```{r, warning=FALSE, error=FALSE}
childcare_plot1<-childcare %>% 
  select(study_year,county_name, mc_infant, mc_toddler, mc_preschool, mfcc_infant, mfcc_toddler,mfcc_preschool)
         
plot_long<-childcare_plot1 %>%
  pivot_longer(!c(study_year,county_name), names_to = "facility", values_to = "Median_cost")

```


Plot 1:

```{r, warning=FALSE, error=FALSE,fig.width=7, fig.height=4}
ggplot(plot_long)+
  geom_smooth(aes(x=study_year,y=Median_cost, color=facility))+
  labs(title="Aggregated weekly Childcare cost in MD based on facility type: 2013-2018",
       subtitle = "Childcare cost has increased during the years for all facility types\n(mc=center based, mfcc=family childcare)",
       caption = "Center based childcare has had the most cost increase during the years")

```
## Using 2 fucntion form our package:

### A
In this section, I want to know the median childcare cost in years for each facility:

```{r, warning=FALSE, error=FALSE,fig.width=7, fig.height=4}
#how much is the average cost of each facility differ from each other 
mtx<-childcare_plot1 %>% select(-c(study_year,county_name)) %>%  as.matrix()
#find the median cost of each facility across years
col.med<-colMedians(mtx)

plot2<- tibble(facility= names(col.med), price= col.med) %>% 
  separate(facility,into = c("facility", "age"),sep = "_") 

ggplot(plot2)+
  geom_col(aes(x=age, y=price, fill=age))+
  facet_wrap(vars(facility))+
  labs(title="Median childcare cost from 2013-2018 in Maryland based on age and facility",
       subtitle = "Chilcare for infants are more expencive than other groups\nboth in mc and mfcc (mc=center based, mfcc=family childcare)",
       caption = "Center based childcare is more expensive than family based childcare")

```

### B
In this section, I would like to know the median childcare cost in each county in from 2013 to2018

```{r,fig.width=7, fig.height=4}
#Find the median childcare cost in each year
row.med<-rowMedians(mtx)

plot3<- tibble(county= childcare_plot1$county_name , year= childcare_plot1$study_year, price= row.med)

ggplot(plot3)+
  geom_point(aes(x=price, y=reorder(county, price), color=as.factor(year)))+
  labs(y="MD County", x="Median childcare price",
       title = "Median weekly childcare price: 2013 to 2018 in each county in MD",
       subtitle = "Montgomery county and Howard county have the highest weekly childcare cost",
       caption="Childcare cost for all ages",
       color="year")
```


## Use functions from purrr package to learn more about occupations in counties

I only focused on management occupation as one of the highest payed jobs and wanted to see if there is a differenc between the percentage of males and females in management in any of the counties 



```{r, warning=FALSE, error=FALSE}
#What is the occupation of the parents?
childcare3<-childcare %>% filter(study_year>=2016) %>% 
  select(study_year,county_name, 39:53) 


childcare3 %>% 
  group_by(county_name) %>% 
  nest() %>% 
  mutate(Female_managment = map_dbl(data, ~ mean(.x$femp_m)),
         male_managment = map_dbl(data, ~ mean(.x$memp_m))) %>%
  arrange(desc(Female_managment))
  
```
Howard County has the highest percentage of male and female in management.Then Montgomery county and Fredrick county.

```{r, warning=FALSE, error=FALSE}
childcare3 %>% filter(study_year==2018) %>% 
  mutate(managment = map2_dbl(femp_m,memp_m,mean)) %>%
  arrange(desc(managment)) %>% 
  select(study_year,county_name,managment)
```

Families might not be traditional (one man and women), so we can look at the average of these two for better analysis.
On average the counties with highest percentage of individuals in management are Howard county, Montgomery County, and Anne Arundel County

# Conclusion:

In this analysis, I was able to look at the childcare costs in Maryland in years form 2013 to 2018 across different counties. Center based childcare for infants is the most expensive childcare type and age group across years and counties. Furthermore, Howard county and Montgomery county have the highest childcare cost and about 60% of parent work in management (fairely high pay job) which makes it more affordable.



# Functions used:
inner_join(), select(), pivot_longer(), filter(), tifyr::drop_na(), separate(), mutate(), arrange()

### ggplot functions
geom_smooth(),geom_col(), geom_point()

### purrr funcitons
map_dbl(),map2_dbl

